---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

The useful packages:
library(dplyr)
```{r , results="hide"}
library(dplyr)
library(ggplot2)
```

```{r echo = FALSE}
knitr::opts_chunk$set(echo =TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

```{r}
data = read.csv("activity.csv", sep=",", head=T)
attach(data)
str(data)
```
Convert date to Date varibale
```{r convert}
data$date =  as.Date(data$date, format="%Y-%m-%d")
str(data)
summary(data)
```

## What is mean total number of steps taken per day
Calculate the total number of steps taken per day

```{r}
step.total=tapply(data$steps, data$date, sum, na.rm=T)
```

Make a histogram of the total number of steps taken each day

```{r}
qplot(step.total, binwidth=500, xlab="total number of steps of a day")
```


Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(step.total)
median(step.total)
```

## What is the average daily activity pattern?

```{r}
step.average=aggregate(x=list(steps=data$steps), by=list(interval = data$interval), mean, na.rm=T)
#plot(step.average, type="l", xlab="Time interval", ylab= "Average number of steps")
ggplot(step.average, aes(x=interval,y=steps)) + geom_line(colour="blue") + xlab("Time interval") + ylab ("Average number of steps")

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
which(step.average$steps ==max(step.average$steps))
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
total.na = is.na(data$steps)
table(total.na)
```

The mean for that 5-minute interval for filling in all of the missing values in the dataset. 

```{r}
for(i in 1: nrow(data)){
        if (is.na(data$steps[i])){
               j =i %% 288  
                if (j != 0){
                        data$steps[i] = step.average$steps[i %% 288 ] 
                       # print(step.average$steps[i] )
                }else{
                        data$steps[i] = step.average$steps[288]   
                }
               
        } 
}
table(is.na(data$steps))
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
data.new= data.frame(data$steps, data[,2],data[,3])
colnames(data.new) = colnames(data)
total.date=tapply(data.new$steps, data.new$date, sum, na.rm=T)
qplot(total.date, binwidth=500, xlab="total number of steps of a day")
```


```{r}
mean(total.date)
median(total.date)
```

## Are there differences in activity patterns between weekdays and weekends?

Convert date to as.Date
```{r}
data.new$date = as.Date(data.new$date, format= "%Y-%m-%d")
data.new$day = rep("0", nrow(data.new))
for (i in 1 : nrow(data.new)){
        week.day = c("Monday","Tuesday","Wednesday","Thursday","Friday")
        if(weekdays(data.new$date[i]) %in% week.day == T){
                 data.new$day[i] = "weekday"
        }else{
                data.new$day[i] = "weekend"
        }
}
```

```{r}
average.new= aggregate(steps ~ interval + day, data=data.new, mean)
ggplot(average.new, aes(x=interval, y= steps)) + geom_line() + facet_grid(day~.)+xlab("5-minute interval") + ylab("Number of steps")
```





